---
title: "hw8"
author: "Rongzhao Huang"
Due: "11:59pm Saturday 8/6/2016.  Instructors stop answering questions at 5pm."
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
library(dplyr)
library(ggplot2)
library(purrr)
library(stringr)
library(nycflights13)
library(RSQLite)
library(animation)
library(tidyr)
```

### purrr

In this exercise, you'll be using functional programming ideas to simulate the central limit theorem on coin flips.  Using neither loops nor the apply functions, write a simulation pipeline that:

* Generates 1000 samples of each 1, 2, ..., 99, 100 fair coin flips.  So for example, you would simulate flipping 1 coins 1000 times, 2 coins 1000 times, etc.
* For each of the 1, 2, ..., 100 flips, produces a histogram of the proportion of heads.  To be explicit: you will create a histogram for the 1000 proportions calculated from the 1 coin flips.  Then you'll do the same for 2 coin flips, 3, etc.

In your plots
* The x-axis should be limited to 0-1
* The y-axis should be limited to 0-15


```{r, eval = FALSE}
coinFlip <- function(x){
  face = sample(c(1, 0), size = x, replace = TRUE, prob = c(0.5, 0.5))
  return(round(sum(face)/x, digits = 10))
}

coinFlipSimulation <- function(x){
  replicate(1000,coinFlip(x))
}


coinsList <- c(1:100)
set.seed(123456)
simulationResult <- coinsList %>% map(coinFlipSimulation)
result <- as.data.frame(simulationResult)
names(result) <- c(1:100)

save_histogram <- function(i){
 plt = ggplot() + 
       aes(x = result[,i], y = ..density..) +
       geom_histogram(bins = 30) + 
       scale_x_continuous(breaks = seq(0,1,by = 0.1)) +
       scale_y_continuous(breaks = seq(0,15,by = 1)) + 
       labs(x = "proportion of heads",
            y = "counts",
            title = "simulate the central limit theorem on coin flips")
 ggsave(plt,file = str_c("plot", i, ".png", sep = ""), path="~/", device = "png")
}

map(c(1:100),save_histogram)
```

Now that you have all the static images, you could "animate" the progression by turning the sequence of images into a gif.  This can be achieved with the `animation` package in R or the standalone `ImageMagick` utility.  This is just an FYI, we will not be looking for a GIF.  Also, Please do NOT upload the images to GitHub nor bCourses.  We'll just be checking your code.

## SQL

These exercises are meant to help you wrap your head around SQL.  Answer the questions below using SQL queries.  You can check your answer using `dplyr` methods, but your grade will be based on the SQL commands.  The `nycflights13` data set has been put into a SQLlite database for these problems.  For documentation about the variable names, see

```{r, eval = FALSE}
help(package = "nycflights13")
```

```{r}
flights <- src_sqlite("nycflights13.mysqlite3")
View(flights)
```


1. Find the full name of each airline that flew to Dallas Fort-Worth (DFW).
```{r}
Airlines =  flights %>% tbl(sql("
  select distinct
    nyairlines.name
  from 
    nyairlines LEFT JOIN nyflights
    ON nyairlines.carrier = nyflights.carrier
  where
    nyflights.dest LIKE 'DFW'
")) %>%
  collect

Airlines
```

2. Make a table containing the tail number, year of manufacture, model, number of engines, and number of seats of the planes flown by United Airlines (UA).  Sort the results by year manufactured.

```{r}
tbl = flights %>% tbl(sql("
  select distinct
    nyflights.tailnum,
    nyplanes.year,
    nyplanes.model,
    nyplanes.engines,
    nyplanes.seats
  from 
    nyplanes,
    nyflights
  where
    nyflights.carrier = 'UA' AND nyflights.tailnum = nyplanes.tailnum
  order by
    nyplanes.year DESC 
  ")) %>%
  collect

tbl
```

3. Make a table with the following information: full airport name in a column called `Airport` and number of flights from New York City to those airports in 2013 in a column called `NumberOfFlights`.  Sort the results in descending order by flight counts.
```{r}
tbl2 = flights %>% tbl(sql("
  select
    nyairports.name as Airport,
    count(nyairports.name) as NumberOfFlights
  from 
    nyairports LEFT JOIN nyflights
    ON nyairports.faa = nyflights.dest
  where
    nyflights.year = 2013
  group by
    nyairports.name
  order by
    count(nyairports.name) DESC
  ")) %>%
  collect

tbl2
```

4. Find the wind speed during the hour of the sceduled departure time of every flight that had a departure delay of more than 30 minutes.  For example, if a flight was scheduled to leave at 5:59am, you would check the wind speed for 5:00am.  In practice, you would of course be more careful than this.
```{r}
tbl3 <- flights %>% tbl(sql("
   select
    flight,
    sched_dep_time,
    dep_time,
    nyweather.wind_speed
    
  from 
    nyflights,
    nyweather
  where
    nyflights.dep_delay > 30 AND nyflights.origin = nyweather.origin AND nyflights.time_hour = nyweather.time_hour 
  ")) %>%
  collect
tbl3
```