---
title: "lab15"
author: "Rongzhao Huang"
date: "July 4, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(xml2)
```

## Example of Checking Work
The best way to deal with the `xml` data is to copy and paste it into a plain-text editor and save it as a `.xml` file.  Windows users, it's highly suggested you either use a Cygwin utility like nano or download Notepad++ rather than use Notepad.  I've saved the following data into `weather.xml` in the lab folder:

```{r, eval = FALSE}
<?xml version="1.0" encoding="UTF-8"?>
<current_observation>

<credit>NOAA's National Weather Service</credit>
<credit_URL>http://weather.gov/</credit_URL>

<image>
  <url>http://weather.gov/images/xml_logo.gif</url>
  <title>NOAA's National Weather Service</title>
  <link>http://weather.gov</link>
</image>

<location>New York/John F. Kennedy Intl Airport, NY</location>
<station_id>KJFK</station_id>
<latitude>40.66</latitude>
<longitude>-73.78</longitude>
<observation_time_rfc822>Mon, 11 Feb 2008 06:51:00 -0500 EST
</observation_time_rfc822>

<weather>A Few Clouds</weather>
<temp_f>11</temp_f>
<temp_c>-12</temp_c>
<relative_humidity>36</relative_humidity>
<wind_dir>West</wind_dir>
<wind_degrees>280</wind_degrees>
<wind_mph>18.4</wind_mph>
<wind_gust_mph>29</wind_gust_mph>
<pressure_mb>1023.6</pressure_mb>
<pressure_in>30.23</pressure_in>
<dewpoint_f>-11</dewpoint_f>
<dewpoint_c>-24</dewpoint_c>
<windchill_f>-7</windchill_f>
<windchill_c>-22</windchill_c>
<visibility_mi>10.00</visibility_mi>

<icon_url_base>http://weather.gov/weather/images/fcicons/</icon_url_base>
<icon_url_name>nfew.jpg</icon_url_name>
<disclaimer_url>http://weather.gov/disclaimer.html</disclaimer_url>
<copyright_url>http://weather.gov/disclaimer.html</copyright_url>

</current_observation>
```

To read the data, we'll need `read_xml()` from the `xml2` package.
```{r}
library(xml2)
weather <- read_xml("weather.xml")
```

To test your XPath, use `xml_find_all()`.  The `text()` at the end grabs the text nodes.
```{r, eval = F}
xml_find_all(xml_doc, xpath)
xml_find_all(weather, "/current_observation/latitude/text()")
xml_find_all(weather, "//latitude/text()")
```

We'll go over the package in more detail tomorrow.  Today, let's just focus on understanding XML structure and XPath syntax.

## Basic XPath

Write XPaths that:

1. Get the contents of the attribute `name`.
```{r}
f = read_xml("lab15_1.xml")
xml_attr(f, "name")

```
2. Select all document elements beneath the node `linkList` with the name `A`.
```{r}
f = read_xml("lab15_2.xml")
xml_find_all(f, "//linkList[@name = 'A']")
```

3. Select all jobs with `priority` corresponding to "critical" or "high".
```{r}
f = read_xml("lab15_3.xml")
xml_find_all(f, "//job[@priority = 'critical' or @priority = 'high']" )
```

4. Select all persons older than 25
```{r}
f = read_xml("lab15_4.xml")
xml_find_all(f, "//person[@age > 25]")
```

5. Select the names of the artists of every CD.

```{r}
f = read_xml("lab15_5.xml")
xml_find_all(f, "/collection/artist/name")
```
## XPathing

The next five questions refer to the document below.  You may want to draw out the document tree.  Experiment! Do comparisons like =, <, etc. need strings or values? Write XPaths that:

```{r}
f = read_xml("lab15-p2.xml")
```

6.  Select the prices of all the books.
```{r}
xml_find_all(f, "//price")
```
7.  Select books that are available on the web or on kindles.
```{r}
xml_find_all(f, '/bookstore/book[@category="web" or @category = "kindle"]')
```
8.  Select books that have page numbers available.
```{r}
xml_find_all(f, "//pages/..")
```
9.  Select the titles of books priced under $20.
```{r}
xml_find_all(f, "//title[../price<20]")
```
10. Select the authors of books that aren't written in English.
```{r}
xml_find_all(f, "//author[../title[@lang != 'en']]")
```
