---
title: "BML Simulation"
date: "July 25, 2016"
output: html_document
---

## Question

* Is there a largest traffic density that permits free flow with no traffic jams?
* Is there a density above which traffic jams are inevitable?

---

## BML Model

* The BML model is essentially a simplified model of traffic flow in a large city grid.

* Each square in a grid of streets contains either a red car, a blue, car, or an empty space.

* At each odd-numbered time step, all red cars simultaneously attempt to move one unit east.

  + A car succeeds if there is an empty space for it to move into.
  
  + If the square it's attempting to move into is occupied, then the car    will remain in its original position.
  
* At each even-numbered time step, all blue cars attempt to move one unit north in the same manner.

* A red car that falls off the east edge of the grid reappears in the same position on the west edge.

* Similarly, a blue car that falls off the north edge of the grid should reappear in the same position on the south edge.

---

## Simulating the BML Model in R

* Represent the grid of streets with an mxn matrix  

* The matrix can only contain the entries: 0 (no cars), 1 (red cars), or 2 (blue cars)

* Distribute the cars at random: each square is independently assigned a car with probability p, or an empty space with probability 1-p

* Each car is equally likely to be red or blue

---

## Matrices in R

* Matrices are a rectangular collection of elements
* Can only contain homogeneous elements (e.g. all numeric, all character, or all logical)

`matrix(data, nrow, ncol, byrow = FALSE)` takes in a vector of data values and creates a matrix of specified dimensions

```{r}
m <- matrix(1:6, nrow = 2, ncol = 3, byrow = FALSE)
m
m <- matrix(1:6, nrow = 2, ncol = 3, byrow = TRUE)
m
```

Assign names to the rows and columns of a matrix with `rownames()` and `colnames()`

```{r}
colnames(m) <- letters[1:3]
m
```

Find the dimensions of a matrix using `dim()`

```{r}
dim(m)
```

Exchange rows and columns using `t()`

```{r}
t(m)
```

Extract out elements of a matrix using the same methods as you would for data frames
```{r}
m[-1, c(2, 3)]
m[ ,"a"]
m[c(TRUE, FALSE), ]
```

Note: `*` multiplies matrices element by element, rather than according to rules of matrix multiplication

```{r}
a <- matrix(1:6, nrow = 2, ncol = 3)
a
b <- matrix(c(0,1,1,0,1,0), nrow = 2, ncol = 3)
b
a*b
```




