---
title: "lab25"
date: "July 26, 2016"
output: html_document
---

Write a function called `moveCars()` that moves the system one full step; that is, the function should move the red cars east once and the blue cars north once, with red cars going first.

Your function should take the following inputs:
`m`: a matrix

Your function should return a list containing two elements:
`m`: an updated matrix
`not_gridlocked`: a logical value that should be TRUE if the system changed and FALSE if not

```{r}
moved <- FALSE

moveCars <- function(m) {
  blocked <- FALSE
  #red cars
  for (row in 1:nrow(m)) {
    for (column in ncol(m):1) {
      
        if (m[row, column] == 1) {
          
          if (column + 1 > ncol(m)) {
              new_col <- 1
              if (m[row, new_col] == 0) {
                blocked <- TRUE
              }
        
        } else {
          new_col <- column + 1
        }
        
          if ((column == 1 & blocked == FALSE & m[row, new_col] == 0) | 
              (column != 1 & m[row, new_col] == 0)) {
            m[row, new_col] <- m[row, column]
            m[row, column] <- 0
            moved <- TRUE
          }
      }
    }
  }
  blocked <- FALSE
  
  #blue cars
  for (row in 1:nrow(m)) {
    for (column in 1:ncol(m)) {
      
      if (m[row, column] == 2) {
          
            if (row - 1 < 1) {
            new_row <- nrow(m)
            if (m[new_row, column] == 0) {
              blocked <- TRUE
            }
        
      } else {
          new_row <- row - 1
        }
        
        if ((row == nrow(m) & blocked == FALSE & m[new_row, column] == 0) |
            (row != nrow(m) & m[new_row, column] == 0)) {
          m[new_row, column] <- m[row, column]
          m[row, column] <- 0
          moved <- TRUE
        }
      }
    }
  }
  x <- list(m, moved)
  return(x)
}
```


Check:
```{r}
m <- matrix(c(1,0,0,0,2,1,2,0,0,2,0,1,1,0,1,2,0,1,2,1,0,0,1,1,2), nrow = 5, ncol = 5, byrow = T)
identical(moveCars(m), list(matrix(c(0,1,0,0,2,1,2,0,0,2,1,1,0,1,0,2,0,1,2,1,0,0,1,1,2), nrow = 5, ncol = 5, byrow = T), TRUE))

m <-  matrix(c(0,1,0,2,1), nrow = 1, ncol = 5, byrow = T)
identical(moveCars(m), list(matrix(c(1,0,1,2,0), nrow = 1, ncol = 5, byrow = T), TRUE))

m <- matrix(c(2,0,1,2,0), nrow = 5, ncol = 1, byrow = T)
identical(moveCars(m), list(matrix(c(0,0,1,2,2), nrow = 5, ncol = 1, byrow = T), TRUE))
```

Now write a function called `initGrid()` that randomizes the initial system.

Your function should take the following inputs:
`r`: number of rows in the matrix
`c`: number of columns in the matrix
`d`: density of the grid 

Your function should return the following:
`m`: an rxc matrix with entries 0 (no cars), 1 (red cars), or 2 (blue cars)

```{r}
initGrid <- function(r, c, d) {
  m <- matrix(NA, nrow = r, ncol = c, byrow = FALSE)
  for (c in 1:ncol(m)) {
    for(r in 1:nrow(m)) {
      m[r, c] <- sample(c(0:1), size = 1, replace = TRUE, prob = c(d, 1 - d))
      if (m[r, c] != 0) {
        m[r, c] <- sample(c(1:2), size = 1, replace = TRUE, prob = c(0.5, 0.5))

      }
    }
  }
  return(m)
}

initGrid(5, 5, 0.4)
```






