# lab3
Read in csv file:
```{r}
library(readr)
tents <- read_csv("tents.csv", col_types = "ccnniccc")

```
1. It tells us the column names of the data frame and the type of each column, plus a few examples in which column.
```{r}
str(tents)
```

2. **tents** is a tbl_df, table, dataframe objects. 

3. rows in **tents**
```{r}
nrow(tents)
```

4. colums in **tents**
```{r}
ncol(tents)
```

5. names of colomns:
```{r}
colnames(tents)
```

6. the first 5 rows as follows
```{r} 
head(tents, 5)
```

7. the last 4 rows as follows
```{r}
tail(tents, 4)
```

8. 75th percentile - 50th percentile of the weights of the tents.
```{r}
quantile(tents$weight)[3:4]
```

9. use n-1 as denominator
```{r}
sd(tents$height)
```

10. adding weight_lbs colomn
```{r}
tents$weight_lbs = tents$weight * 0.0022046
```

11. adding height_in colomn
```{r}
tents$height_in = tents$height * 0.393701
colnames(tents)
```

12. number of tents cost more than $400
```{r}
length(tents$price[tents$price > 400])
```

13. subset tents into so_cheap_tents
```{r}
sp_cheap_tents = subset(tents, tents$price <= 300)
sp_cheap_tents
```

14. The percentage of big-agnes within the group > 300
```{r}
total = subset(tents, tents$price > 300)
denom = nrow(total)
nom = nrow(subset(total, total$brand == "big-agnes"))
percentage = nom / denom
percentage
```

15. Select the bestuse in 4 different ways
```{r}
tents$bestuse
tents[6]$bestuse
tents[["bestuse"]]
tents[[6]]
tents["bestuse"]$bestuse
```

16. The best height-weight ratio tent
```{r}
tents$name[which.max(tents$height/tents$weight)]
```

17. It creates a table with seasons being row indice and capacity being colomn indice
```{r}
table(tents$seasons, tents$capacity)
```

18. The most common type of tent is
```{r}
names(sort(table(tents$bestuse), decreasing = TRUE))[1]
```

19. Create Kelty, then Big-Agnes, then combine.
```{r}
kelty = subset(tents, tents$brand == "kelty")
big_agnes = subset(tents, tents$brand == "big-agnes")
rbind(kelty, big_agnes)
```