---
title: "lab6"
author: "Andrew Do"
date: "June 15, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Population Variance

Write a function `MyVar` that calculates the population variance of a vector.  If you need a refresher of what that means, [this section of wikipedia should help](https://en.wikipedia.org/wiki/Variance#Population_variance_and_sample_variance).

```{r}
MyVar = function(x){
  var(x)
}
```

## Finding Roots
Write a function `zeroes` that finds the roots (zeroes) of a degree two polynomial.  The function should take the coefficients of the polynomial $a^2x + bx + c$ as its arguments and returns a vector of roots.  If there are no real roots, the function should stop and produce an error message: `"This polynomial has no real roots."`

```{r}
zeroes = function(a, b, c){
  if((b^2 - 4*a*c) < 0){
    stop("This polynomial has no real roots.")}
  else if((b^2 - 4*a*c) == 0){
    root = (-b / (2*a))
    return(root)
  }
  else{
    root = c((-b + sqrt(b^2 - 4*a*c))/(2*a), (-b - sqrt(b^2 - 4*a*c))/(2*a)  )
    return(root)
  }
}
```

Test your function:
```{r, eval = F}
zeroes(1, 0, 0)
zeroes(1, 0, 1)
zeroes(1, 1, 1)
zeroes(1, 6, 5)
```

## Temperature Conversion

Write a function `ConvertTemperature` that takes a numeric vector of temperatures in Fahrenheit and converts them into Celsius, Kelvin, Rankine, Delisle, Newton, Reamur, or Romer, depending on what the user specifies in an argument called `to`.  If they specify some other temperature scale, return the message "The [their input] temperature scale is not supported".  If you don't know the conversions, (see this table)[https://en.wikipedia.org/wiki/Conversion_of_units_of_temperature#Fahrenheit]

```{r}
ConvertTemperature = function(degree, unit){
  if(unit == "Celsius" | unit == "celsius"){
    ret = (degree - 32) * (5/9)
    return(round(ret, digit = 2))
  }
  else if(unit == "Kelvin" | unit == "kelvin"){
    ret = (degree + 459.67) *(5/9)
    return(round(ret, digit = 2))}
  else if(unit == "Rankine" | unit == "rankine"){
    ret = degree + 459.67
    return(round(ret, digit = 2))
  }
  else{
    stop("we couldn't convert into this unit.")
  }
}
```

Test your function
```{r}
ConvertTemperature(32, "Celsius") == 0
ConvertTemperature(212, "Kelvin") == 373.15
ConvertTemperature(10, "Rankine") == 469.67
```

## Rolling Dice

For this problem, you'll need the `sample` and `pmax` function.  `sample` generates a random sample of a specified size either with or without replacement.  You can specify probability weights for each element of your sample space using the `prob` argument.  By default, each element will have equal probability.
```
sample(x = vector_to_be_sampled, size, replace = FALSE, prob = NULL)
```
1. What does `pmax` do? How is it different from `max`? Hint: test both on the two vectors `c(1, 5, 3)` and `c(2, 4, 2)`.

**Answer: pmax returns a vector of greater number with the same index within two vectors. max picks out the greatest number in all the vectors.**

2. When you "roll dice with advantage", you roll a pair of dice and take the higher value as your roll.  Write a function `AdvRoll` the simulates rolling with advantage `n` times. The function should take on 2 arguments `sides` and `n` and output a length `n` vector of rolls.

```{r}
AdvRoll = function(sides, n){
  ret = seq(n)
  for(i in ret){
    roll_1 = sample(seq(sides), 1, replace = TRUE, prob = NULL)
    roll_2 =  sample(seq(sides), 1, replace = TRUE, prob = NULL)
    num = max(roll_1, roll_2)
    ret[i] = num
  }
  return(ret)
}
```

3. Write your own summary function called `MySummary` that, given a numeric vector `x`, returns the following summary statistics: mean, median, mode, sum, and sum of squares.

```{r}
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
MySummary = function(x){
  sprintf("mean: %f, median: %f, mode: %f, sum: %f, sum of squares: %f", mean(x), median(x), getmode(x), sum(x), sum(x^2))
}
MySummary(c(1,2,3,4,5, 6, 7, 8, 9, 10))
```

4. Using your two functions, simulate "rolling dice with advantage" 1000 times and find the summary statistics.

```{r}
MySummary(AdvRoll(6, 1000))
```