---
title: "lab5"
author: "Andrew Do"
date: "June 14, 2016"
output: html_document
---

```{r setup, include=FALSE}
library(stringr)
library(readr)
knitr::opts_chunk$set(echo = TRUE)
```

## Getting Help for Packages

You can see what an entire package has to offer with
```{r, eval = F}
help(package = "stringr")
```


## Counting number of vowels

Consider the following vector `letrs` which contains various letters:
```{r}
# Garbage string
letrs <- "gfngenodgsyctyviqyyigeinyngbkwjdjqkrrollhpxdkfldlytympmcgmjynihusebtq"
```

If you were to count the number of vowels in `letrs` you would get the following counts:

- a: 0
- e: 3
- i: 4
- o: 2
- u: 1

This exercise will guide you through how to arrive at this frequency table.

```{r}
# Create a character vector of vowels.  Give it an appropriate name.
x = c()
# Use str_count to get the counts.  Assign this an appropriate name.
x = str_count(letrs, c("a", "e", "i", "o", "u"))

# Give each element of the vector a name using the names function.
# The result should look like:
# a e i o u
# 0 3 4 2 1
names(x) = c("a", "e", "i", "o", "u")
x
```

## Exploring new functions: str_pad
Write a short description of what each of the following calls of `str_pad` is doing.  The last three examples were found at the bottom of the function's documentation.

1. In this call of `str_pad`, creates a padding of x repeated 4 times starting from the left, concatenate the padding with an 'a'.

```{r}
str_pad("a", width = 4, side = "left", pad = "x")
```

2. In this call of `str_pad`, as same as above other than the default padding is empty space.

```{r}
str_pad("a", width = 3, side = "left")
```

3. In this call of `str_pad`, it applies the str_pad function to each element of the character vector, with default padding being empty space. Starting from the left, the padding ending with the chacrater fills total of 10 digit width.

```{r}
str_pad(c("a", "abc", "abcdef"), 10)
```

4. In this call of `str_pad`, it applies the empty space padding to 'a' three times from the left, with total width being 5, 10, and 20.

```{r}
str_pad("a", c(5, 10, 20))
```

5. In this call of `str_pad`, it applies the empty space padding to 'a' three times from the left, with total width being 10, but each time with different padding, -, _, and empty spance.

```{r}
str_pad("a", 10, pad = c("-", "_", " "))
```

## Regular Expressions

We read in the data from the csv file and extract the names column
```{r}
flags <- read_csv("flags.csv")
countries <- flags$name
```

You may have noticed that in the `countries` vector, words in a country's name are separated by `-`.  Use `str_replace_all` to replace the hyphens with spaces.  For example `"Vatican-City"` should be replaced with `"Vatican City"`

```{r}
str_replace_all(countries, '-', ' ')
```

Use `str_detect` to see if there are any countries with the world "island" in them, ignoring case.  You may also want to look up how to use `ignore.case`.

```{r}
any(str_detect(countries, ignore.case('island')))
```

Use `str_subset` to find all countries that begin with a capital `A`.
```{r}
str_subset(countries, '^A')
```

Find all countries that end with `stan`
```{r}
str_subset(countries, 'stan$')
```

Use `str_replace_all` to replace the `$`'s in `baws` with `s`'s.  Recall that the `$` is a metacharacter.
```{r}
baws <- "Bo$$"
str_replace_all(baws, '\\$', 's')
```

Use a character class (denoted by square brackets) to find all countries beginning with a vowel and ending with a consonant.
```{r}
vowel_class = "^[AEIOUaeiou]"
intermediate = str_subset(countries, vowel_class)
consonant_class = "[^aeiouAEIOU]$"
str_subset(intermediate, consonant_class)
```




